Bacterial Microbiome and Obesity Meta-analysis Statistics Pipeline

### 2016-02-06 code/StatisticsProcessing.Rmd
twinsUK update

##### **Goodrich et al. 2014:**  
```{r}

@@

StudyNo <- c(1, 2, 3, 4, 5, 6, 7, 8)
Study <- c("Baxter", "HMP", "Ross", "Escobar", "Arumugam", "Turnbaugh", "Zupancic", "Wu")
Year <- c("2015", "2011", "2015", "2014", "2010", "2009", "2012", "2012")
Total <- c("n=172", "n=256", "n=63", "n=30", "n=85", "n=146", "n=207", "n=58")
tpos <- c(25, 16, 20, 6, 20, 52, 47, 2) #obese in low diversity group
tneg <- c(61, 112, 12, 9, 23, 21, 57, 27) #nonobese in low diversity group
cpos <- c(22, 10, 18, 4, 17, 47, 31, 3) #obese in high diversity group
cneg <- c(64, 118, 13, 11, 25, 26, 72, 26) #nonobese in high diversity group
RR <- c(1.14, 1.6, 1.08, 1.50, 1.15, 1.11, 1.50, 0.67)
low <- c(0.70, 0.75, 0.72, 0.53, 0.71, 0.88, 1.04, 0.12)
high <- c(1.85, 3.39, 1.61, 4.26, 1.87, 1.38, 2.16, 3.70)

@@

# RR Title ################################################
#text(3.2, 3.10, "Relative Risk [95% CI]", pos=2, cex=0.8)#
  #large scale (Plot in R studio at full)                 #
                                                          #
text(5.5, 9.3, "Relative Risk [95% CI]", pos=2, cex=0.8)  #
  #smaller scale (Plot in R studia at normal)             #
###########################################################

# Group Title#############################################
                                                         #
text(-6.5, 9.3, "Study, Year, and n", pos=4, cex=0.8)    #
                                                         #
##########################################################

@@

##Calculate combined P-value
m = 8
T = -2*log(0.608)-2*log(0.214)-2*log(0.236)-2*log(0.439)-2*log(0.575)-2*log(0.376)-2*log(0.025)-2*log(0.64)
T.pval = 1-pchisq(T,m) #P-value = 0.01

##Alternatively can use built in package (I trust this more)
anova.rma(res)
# QM = 5.1101, p-value = 0.0238

@@

StudyNo <- c(1, 2, 3, 4, 5, 6, 7, 8)
Study <- c("Baxter", "HMP", "Ross", "Escobar", "Arumugam", "Turnbaugh", "Zupancic", "Wu")
Year <- c("2015", "2011", "2015", "2014", "2010", "2009", "2012", "2012")
Total <- c("n=172", "n=256", "n=63", "n=30", "n=85", "n=146", "n=207", "n=58")
tpos <- c(24, 13, 16, 3, 17, 45, 37, 2) #obese in low BF group
tneg <- c(62, 115, 16, 12, 26, 28, 67, 27) #nonobese in low BF group
cpos <- c(23, 13, 22, 7, 20, 54, 41, 3) #obese in high BF group
cneg <- c(63, 115, 9, 8, 22, 19, 62, 26) #nonobese in high BF group
RR <- c(1.04, 1.00, 0.70, 0.43, 0.83, 0.83, 0.89, 0.67)
low <- c(0.64, 0.48, 0.47, 0.14, 0.51, 0.66, 0.66, 0.12)
high <- c(1.70, 2.07, 1.07, 1.35, 1.35, 1.05, 1.27, 3.70)

@@

# RR Title ################################################
#text(3.2, 3.10, "Relative Risk [95% CI]", pos=2, cex=0.8)#
#large scale (Plot in R studio at full)                 #
#
text(5.5, 9.3, "Relative Risk [95% CI]", pos=2, cex=0.8)  #
#smaller scale (Plot in R studia at normal)             #
###########################################################

# Group Title#############################################
#
text(-6.0, 9.3, "Study, Year, and n", pos=4, cex=0.8)    #
#
##########################################################

@@

##Calculate combined P-value
m = 8
T = -2*log(0.864)-2*log(1.00)-2*log(0.089)-2*log(0.121)-2*log(0.452)-2*log(0.111)-2*log(0.53)-2*log(0.64)
T.pval = 1-pchisq(T,m) #P-value = 0.03155827

##Alternatively can use built in package (I trust this more)
anova.rma(res)
  # QM = 5.3494, p-value = 0.0207

@@

#Amish STD = 0.5893173
amish.ZScore.H <- amish.data$H.corr/0.5893173
hist(amish.ZScore.H)
#Turnbaugh STD = 0.4818751
turn.ZScore.H <- turn.data$H.corr/0.4818751
hist(turn.ZScore.H)
#MetaHit-Den STD = 0.4038558
keepmetahit <- which(data1$dataset == "MetaHit")
metahitdata <- data1[keepmetahit, ]
metahit.ZScore.H <- metahitdata$H.corr/0.4038558
hist(metahit.ZScore.H)
#Korean STD = 0.5400615
keepKorean <- which(data1$dataset == "Korean")
Koreandata <- data1[keepKorean, ]
Korean.ZScore.H <- Koreandata$H.corr/0.5400615
hist(Korean.ZScore.H)
#HMP STD = 0.6774514
keepHMP <- which(data1$dataset == "HMP")
hmpdata <- data1[keepHMP, ]
HMP.ZScore.H <- hmpdata$H.corr/0.6774514
hist(HMP.ZScore.H)
#CCHC STD = 0.5784522
keepCCHC <- which(data1$dataset == "CCHC")
CCHCdata <- data1[keepCCHC, ]
CCHC.ZScore.H <- CCHCdata$H.corr/0.5784522
hist(CCHC.ZScore.H)
#GLNE07 STD = 0.4663871
keepGLNE <- which(data1$dataset == "GLNE")
GLNEdata <- data1[keepGLNE, ]
GLNE.ZScore.H <- GLNEdata$H.corr/0.4663871
hist(GLNE.ZScore.H)
#Escobar STD = 0.5378935
keepescobar <- which(data1$dataset == "Escobar")
escobardata <- data1[keepescobar, ]
Escobar.ZScore.H <- escobardata$H.corr/0.5378935
hist(Escobar.ZScore.H)
#COMBO STD = 0.6870763
keepCOMBO <- which(data1$dataset == "COMBO")
COMBOdata <- data1[keepCOMBO, ]
COMBO.ZScore.H <- COMBOdata$H.corr/0.6870763
hist(COMBO.ZScore.H)

#data1 order Escobar, GLNE, CCHC, HMP, Korean, MetaHit
HZScore <- c(Escobar.ZScore.H, GLNE.ZScore.H, CCHC.ZScore.H, HMP.ZScore.H, Korean.ZScore.H, metahit.ZScore.H, COMBO.ZScore.H)
data1 <- cbind(HZScore, data1)

amish.data <- cbind(amish.ZScore.H, amish.data)
turn.data <- cbind(turn.ZScore.H, turn.data)
rm(Escobar.ZScore.H, GLNE.ZScore.H, CCHC.ZScore.H, HMP.ZScore.H, Korean.ZScore.H, metahit.ZScore.H, amish.ZScore.H, turn.ZScore.H, COMBO.ZScore.H)
rm(keepmetahit, keepCCHC, keepescobar, keepGLNE, keepHMP, keepKorean)

@@

group.data.test1 <- cbind(comb1.H.Zscore, comb1.obese, comb1.bmi.class)
group.data.test2 <- cbind(comb2.H.Zscore, comb2.obese)
colnames(group.data.test1) <- c("H.Zscore", "obese", "BMI.class")
colnames(group.data.test2) <- c("H.Zscore", "obese")
#Look at Obese versus non-obese
t.test(H.Zscore ~ obese, data=group.data.test2) #P-value = 0.002583
table(group.data.test2$obese) # No = 695 | Yes = 340
p1 <- ggplot(group.data.test2, aes(obese, H.Zscore)) + geom_boxplot()

#A more in-depth look, breaking down the groups
anova(lm(H.Zscore ~ BMI.class, data=group.data.test1)) #P-value = 0.01783
a2 <- aov(H.Zscore ~ BMI.class, data=group.data.test1)
posthoc <- TukeyHSD(x=a2, 'BMI.class', conf.level=0.95)
  # Obese-Lean (p-value=0.070), Over-Norm (p-value=0.791), Over-Obese (p-value=0.021)

table(group.data.test1$BMI.class) 
  # Lean = 326 | Overweight = 240 | Obese = 262
ddply(group.data.test1, ~BMI.class, summarise, mean=mean(H.Zscore), sd=sd(H.Zscore))
  # Lean = 0.042 +\- 0.940 | Overweight = 0.097 +/- 0.993 | Obese = -0.140 +\- 1.052

@@

summary(lm(HZScore ~ bmi, data=data1)) #P-value 0.06558, R2= 0.004977
plot(HZScore ~ bmi, data=data1)
abline(lm(HZScore ~ bmi, data=data1), col="red")

@@

#Amish SD = 0.7918527
am.BF.corr <- as.character(amish.data$BF.ratio.corr)
am.BF.corr <- as.numeric(am.BF.corr) # Value 19 has no value
amish.norm.BF <- am.BF.corr/0.7918527
hist(amish.norm.BF)
#Turnbaugh SD = 0.4578937
turn.norm.BF <- turn.data$BF.corr/0.4578937
hist(turn.norm.BF)
#MetaHit SD = 0.9707577
keepmetahit <- which(data2$dataset == "MetaHit")
metahitdata <- data2[keepmetahit, ]
metahit.norm.BF <- metahitdata$BF.ratio.corr/0.9707577
hist(metahit.norm.BF)
#Korean SD = 0.6048195
keepKorean <- which(data2$dataset == "Korean")
Koreandata <- data2[keepKorean, ]
Korean.norm.BF <- Koreandata$BF.ratio.corr/0.6048195
hist(Korean.norm.BF)
#HMP SD = 8.301864
keepHMP <- which(data2$dataset == "HMP")
hmpdata <- data2[keepHMP, ]
HMP.norm.BF <- hmpdata$BF.ratio.corr/8.301864
hist(HMP.norm.BF)
#CCHC SD = 2.00527
keepCCHC <- which(data2$dataset == "CCHC")
CCHCdata <- data2[keepCCHC, ]
CCHC.norm.BF <- CCHCdata$BF.ratio.corr/2.00527
hist(CCHC.norm.BF)
#GLNE07 SD = 0.4451074
keepGLNE <- which(data2$dataset == "GLNE")
GLNEdata <- data2[keepGLNE, ]
GLNE.norm.BF <- GLNEdata$BF.ratio.corr/0.4451074
hist(GLNE.norm.BF)
#Escobar SD = 0.4232106
keepescobar <- which(data2$dataset == "Escobar")
escobardata <- data2[keepescobar, ]
Escobar.norm.BF <- escobardata$BF.ratio.corr/0.4232106
hist(Escobar.norm.BF)
#COMBO SD = 5.928664
keepCOMBO <- which(data2$dataset == "COMBO")
COMBOdata <- data2[keepCOMBO, ]
COMBO.norm.BF <- COMBOdata$BF.ratio.corr/5.928664
hist(COMBO.norm.BF)

#data1 order Escobar, GLNE, CCHC, HMP, Korean, MetaHit
BFScore <- c(Escobar.norm.BF, GLNE.norm.BF, CCHC.norm.BF, HMP.norm.BF, Korean.norm.BF, metahit.norm.BF, COMBO.norm.BF)
data2 <- cbind(BFScore, data2)

amish.data <- cbind(amish.norm.BF, amish.data)
turn.data <- cbind(turn.norm.BF, turn.data)
rm(Escobar.norm.BF, GLNE.norm.BF, CCHC.norm.BF, HMP.norm.BF, Korean.norm.BF, metahit.norm.BF, amish.norm.BF, turn.norm.BF, COMBO.norm.BF)
rm(keepmetahit, keepCCHC, keepescobar, keepGLNE, keepHMP, keepKorean)

@@

#Look at Obese versus non-obese
t.test(BF.ratio.corr ~ obese, data=group.data.test2) 
  # P-value = 0.0525
table(group.data.test2$obese)
  # No = 695 | Yes = 340
bf1 <- ggplot(group.data.test2, aes(obese, BF.ratio.corr)) + geom_boxplot()

#A more in-depth look, breaking down the groups
anova(lm(BF.ratio.corr ~ BMI.class, data=group.data.test)) 
  # P-value = 0.0227
a2 <- aov(BF.ratio.corr ~ BMI.class, data=group.data.test)
posthoc <- TukeyHSD(x=a2, 'BMI.class', conf.level=0.95)
  # Obese-Norm (p-value=0.026) | Over-Norm (p-value=0.117), Over-Obese (p-value=0.863)
table(group.data.test$BMI.class)
  # Lean = 326 | Overweight = 240 | Obese = 262
ddply(group.data.test, ~BMI.class, summarise, mean=mean(BF.ratio.corr), sd=sd(BF.ratio.corr))
  # Lean = -0.361 +\- 1.431 | Overweight = -0.695 +/- 2.032 | Obese = -0.787 +\- 2.461
bf2 <- ggplot(group.data.test, aes(BMI.class, BF.ratio.corr)) + geom_boxplot()

summary(lm(BFScore ~ bmi, data=data2)) #P-value 0.07859, R2= 0.004543
plot(BFScore ~ bmi, data=data2)
abline(lm(BFScore ~ bmi, data=data2), col="red")

@@

###################################Bacteroidetes#######################################
#Look at Obese versus non-obese
t.test(Zbacter ~ obese, data=group.data.test2) 
  # P-value = 0.3202
table(group.data.test2$obese)
  # No = 695 | Yes = 340
bacter1 <- ggplot(group.data.test2, aes(obese, Zbacter)) + geom_boxplot()

>>>>

##### **Goodrich et al. 2014:**  
```{r, eval=FALSE}
###########################################################################
############ Preparing Data Tables for Analysis ###########################
###########################################################################

setwd("C:/users/marc/Desktop/obesity2/twinsUK")

#Read in and match metadata to microbiome data
metadata <- read.csv("TwinsUKStudy_metadata.csv")
rownames(metadata) <- metadata[, 1]
shared.data <- read.table("combined.tx.1.subsample.shared", header=T)
rownames(shared.data) <- shared.data[, 2]
microbiome <- shared.data[, -c(1:3)]
keep  <- rownames(metadata)

microbiome<- microbiome[keep, ]

rm(good.metadata, shared.data, keep)

#generate alpha diversity measures with vegan
library(vegan)
H <- diversity(microbiome)
S <- specnumber(microbiome)
J <- H/log(S)
alpha.diversity <- as.data.frame(cbind(H, S, J))
alpha.test <- alpha.diversity

#Get phyla information
#Edited out non phyla information first with sed in linux
#combined new labels with previous taxonomy file with excel
phylogenetic.info <- read.table("phyla.txt", header=T)
rownames(phylogenetic.info) <- phylogenetic.info[,1]
phylogenetic.info <- phylogenetic.info[,-c(1)]
phyla.names <- as.character(phylogenetic.info$Taxonomy)
keep <- colnames(microbiome)
phyla.good <- phylogenetic.info[keep, ]
phyla.names <- as.character(phyla.good[,2])
phyla.table <- microbiome
colnames(phyla.table) <- phyla.names
#add all the same columns up and then return the sum
testing <- t(rowsum(t(phyla.table), group = rownames(t(phyla.table))))
phyla.table <- as.data.frame(testing)
rm(testing)
#combine phyla that are not that abundant
phyla.table$other <- apply(phyla.table[, c("Acidobacteria", "Chloroflexi", "Deferribacteres", "Elusimicrobia", "Fusobacteria", "Gemmatimonadetes", "Lentisphaerae", "OD1", "Planctomycetes", "Spirochaetes", "SR1", "Synergistetes", "Tenericutes")], 1, sum)
phyla.table <- phyla.table[, -c(1, 4:6, 8:12, 14:17)]

#Create a relative abundance table for phyla
phyla.total <- apply(phyla.table[, c(1:7)], 1, sum)
phyla.table.rel.abund <- (phyla.table/phyla.total)*100

#Create BMI groups
metadata$BMI.class[metadata$body_mass_index_s<=24] <- "Normal"
metadata$BMI.class[metadata$body_mass_index_s>24 & metadata$body_mass_index_s<30] <- "Overweight"
metadata$BMI.class[metadata$body_mass_index_s>=30 & metadata$body_mass_index_s<40] <- "Obese"
metadata$BMI.class[metadata$body_mass_index_s>=40] <- "Extreme Obesity"

#Get paitent demographics to be tested
bmi <- metadata$body_mass_index_s
obese <- factor(metadata$obese)
bmi.class <- factor(metadata$BMI.class)
bmi.class2 <- factor(metadata$BMIclass2)

anova(lm(H ~ obese)) #P-value=0.6671
anova(lm(H ~ bmi.class)) #P-value=0.9732
anova(lm(H ~ bmi.class2)) #P-value=0.8987

anova(lm(S ~ obese)) #P-value=0.9073
anova(lm(S ~ bmi.class)) #P-value=0.9124
anova(lm(S ~ bmi.class2)) #P-value=0.8202

anova(lm(J ~ obese)) #P-value=0.6319
anova(lm(J ~ bmi.class)) #P-value=0.943
anova(lm(J ~ bmi.class2)) #P-value=0.8394

##linear correlations
summary(lm(H ~ bmi)) #P-value=0.8662, R2=5.63x10-5
summary(lm(S ~ bmi)) #P-value=0.9142, R2=2.302x10-5
summary(lm(J ~ bmi)) #P-value=0.8582, R2=6.332x10-5

anova(lm(bacter ~ obese)) #P-value=0.6661
anova(lm(bacter ~ bmi.class)) #P-value=0.3184
anova(lm(bacter ~ bmi.class2)) #P-value=0.3184

anova(lm(firm ~ obese)) #P-value=0.9688
anova(lm(firm ~ bmi.class)) #P-value=0.399
anova(lm(firm ~ bmi.class2)) #P-value=0.2455

anova(lm(BFratio ~ obese)) #P-value=0.8709
anova(lm(BFratio ~ bmi.class)) #P-value=0.7915
anova(lm(BFratio ~ bmi.class2)) #P-value=0.5976

set.seed(3)
adonis(microbiome ~ obese, permutations=1000) 
#PERMANOVA=0.004995, pseudo-F=2.9308

set.seed(3)
adonis(microbiome ~ bmi.class, permutations=1000) 
#PERMANOVA=0.005994, pseudo-F=1.8853

set.seed(3)
adonis(microbiome ~ bmi.class2, permutations=1000) 
#PERMANOVA=0.007992, pseudo-F=2.1453

#Generate median values and put them into existing alpha.test dataframe
#Shannon diversity
median(H) # 2.725646
alpha.test <- within(alpha.test, {shannon.cat = ifelse(H <= median(H), "less", "higher")})

#OTU Richness
median(S) # 73
alpha.test <- within(alpha.test, {S.cat = ifelse(S <= median(S), "less", "higher")})

#Evenness
median(J) # 0.6311328
alpha.test <- within(alpha.test, {J.cat = ifelse(J <= median(J), "less", "higher")})

##Shannon Diversity
H.cat <- alpha.test$shannon.cat
S.cat <- alpha.test$S.cat
J.cat <- alpha.test$J.cat
bmi.cat <- as.character(obese)
test3 <- cbind(H.cat, S.cat, J.cat, bmi.cat)
test3 <- test3[order(H.cat), ]
table(test3[c(1:253), 4])
table(test3[c(254:507), 4])
#Group1 (Higher than median), obese = 47 and non-obese = 206
#Group2 (Lower than median), obese = 56 and non-obese = 198

group1 <- c(47, 206)
group2 <- c(56, 198)
r.test <- rbind(group2, group1)
colnames(r.test) <- c("Obese", "Not.Obese")
rownames(r.test) <- c("group2", "group1")

epi.2by2(r.test, method="cohort.count")
## Risk Ratio = 1.19
## CI = 0.84, 1.68
## p-value = 0.332

median(BFRatio$BFRatio) # 0.5304444
BFRatio <- within(BFRatio, {BFRatio.cat = ifelse(BFRatio <= median(BFRatio), "less", "higher")})

BFRatio.cat <- BFRatio$BFRatio.cat
test4 <- cbind(BFRatio.cat, obese)
test4 <- test4[order(BFRatio.cat), ]
table(test4[c(1:253), 2]) #higher group
table(test4[c(254:507), 2]) #lower group
#Group1 (Higher than median), obese = 52 and non-obese = 201
#Group2 (Lower than median), obese = 51 and non-obese = 203

group1 <- c(52, 201)
group2 <- c(51, 203)
r.test <- rbind(group2, group1)
colnames(r.test) <- c("Obese", "Not.Obese")
rownames(r.test) <- c("group2", "group1")

epi.2by2(r.test, method="cohort.count")
## Risk Ratio = 0.98
## CI = 0.69, 1.38
## p-value = 0.894

#Try AUCRF with default measures provided in readme
set.seed(3)
fit <- AUCRF(obese ~ ., data=testset, ntree=1000, nodesize=20)
summary(fit) # list of 8 Measures, AUCopt = 0.6767159
plot(fit)

#Try AUCRF with only phyla and alpha diversity measures in the testset
testset <- cbind(obese, H, S, J, phyla.table.rel.abund)
set.seed(3)
fit <- AUCRF(obese ~ ., data=testset, ntree=1000, nodesize=20)
summary(fit) 
# list of 7 Measures, AUCopt = 0.538138
# Proteobacteria, unclassified, Verrucomicrobia
plot(fit)
```
<br><br>

@@

StudyNo <- c(1, 2, 3, 4, 5, 6, 7, 8, 9)
Study <- c("Baxter", "HMP", "Ross", "Escobar", "Arumugam", "Turnbaugh", "Zupancic", "Wu", "Goodrich")
Year <- c("2015", "2011", "2015", "2014", "2010", "2009", "2012", "2012", "2014")
Total <- c("n=172", "n=256", "n=63", "n=30", "n=85", "n=146", "n=207", "n=58", "n=507")
tpos <- c(25, 16, 20, 6, 20, 52, 47, 2, 56) #obese in low diversity group
tneg <- c(61, 112, 12, 9, 23, 21, 57, 27, 198) #nonobese in low diversity group
cpos <- c(22, 10, 18, 4, 17, 47, 31, 3, 47) #obese in high diversity group
cneg <- c(64, 118, 13, 11, 25, 26, 72, 26, 206) #nonobese in high diversity group
RR <- c(1.14, 1.6, 1.08, 1.50, 1.15, 1.11, 1.50, 0.67, 1.19)
low <- c(0.70, 0.75, 0.72, 0.53, 0.71, 0.88, 1.04, 0.12, 0.84)
high <- c(1.85, 3.39, 1.61, 4.26, 1.87, 1.38, 2.16, 3.70, 1.68)

@@

# RR Title ################################################
#text(3.2, 3.10, "Relative Risk [95% CI]", pos=2, cex=0.8)#
  #large scale (Plot in R studio at full)                 #
                                                          #
text(5.5, 10.3, "Relative Risk [95% CI]", pos=2, cex=0.8)  #
  #smaller scale (Plot in R studia at normal)             #
###########################################################

# Group Title#############################################
                                                         #
text(-6.5, 10.3, "Study, Year, and n", pos=4, cex=0.8)    #
                                                         #
##########################################################

@@

## Alternative way - 
# Seems easier to follow plus I am not trying to do
# a meta analysis on a model

data <-as.data.frame(cbind(RR, low, high))
raw.overall2 <- as.data.frame(cbind(Study,Year, data))

raw.overall2$logRR <- log(raw.overall2$RR)
raw.overall2$SE <- (log(raw.overall2$high) - log(raw.overall2$low))/(2*1.96)
attach(raw.overall2)
metaplot(logRR, SE, labels=Study, nn=1, logeffect=TRUE, col=meta.colors("black"), xlab="RR of Obesity in high versus low Shannon diversity based on Median", ylab="")
  #Graphical output is not as informative

##Calculate combined P-value
m = 8
T = -2*log(0.608)-2*log(0.214)-2*log(0.236)-2*log(0.439)-2*log(0.575)-2*log(0.376)-2*log(0.025)-2*log(0.64)-2*log(0.332)
T.pval = 1-pchisq(T,m) #P-value = 0.00464

##Alternatively can use built in package (I trust this more)
anova.rma(res)
# QM = 6.0488, p-value = 0.0139

@@

StudyNo <- c(1, 2, 3, 4, 5, 6, 7, 8, 9)
Study <- c("Baxter", "HMP", "Ross", "Escobar", "Arumugam", "Turnbaugh", "Zupancic", "Wu", "Goodrich")
Year <- c("2015", "2011", "2015", "2014", "2010", "2009", "2012", "2012", "2014")
Total <- c("n=172", "n=256", "n=63", "n=30", "n=85", "n=146", "n=201", "n=58", "n=507")
tpos <- c(24, 13, 16, 3, 17, 45, 37, 2, 51) #obese in low BF group
tneg <- c(62, 115, 16, 12, 26, 28, 67, 27, 203) #nonobese in low BF group
cpos <- c(23, 13, 22, 7, 20, 54, 41, 3, 52) #obese in high BF group
cneg <- c(63, 115, 9, 8, 22, 19, 62, 26, 201) #nonobese in high BF group
RR <- c(1.04, 1.00, 0.70, 0.43, 0.83, 0.83, 0.89, 0.67, 0.98)
low <- c(0.64, 0.48, 0.47, 0.14, 0.51, 0.66, 0.66, 0.12, 0.69)
high <- c(1.70, 2.07, 1.07, 1.35, 1.35, 1.05, 1.27, 3.70, 1.38)

@@

# RR Title ################################################
#text(3.2, 3.10, "Relative Risk [95% CI]", pos=2, cex=0.8)#
#large scale (Plot in R studio at full)                 #
#
text(5.5, 10.3, "Relative Risk [95% CI]", pos=2, cex=0.8)  #
#smaller scale (Plot in R studia at normal)             #
###########################################################

# Group Title#############################################
#
text(-6.0, 10.3, "Study, Year, and n", pos=4, cex=0.8)    #
#
##########################################################

@@

##Calculate combined P-value
m = 8
T = -2*log(0.864)-2*log(1.00)-2*log(0.089)-2*log(0.121)-2*log(0.452)-2*log(0.111)-2*log(0.53)-2*log(0.64)-2*log(0.894)
T.pval = 1-pchisq(T,m) #P-value = 0.02338081

##Alternatively can use built in package (I trust this more)
anova.rma(res)
  # QM = 4.7318, p-value = 0.0296

@@

#Amish STD = 0.5893173
amish.ZScore.H <- amish.data$H.corr/0.5893173
hist(amish.ZScore.H)
#Turnbaugh STD = 0.4818751
turn.ZScore.H <- turn.data$H.corr/0.4818751
hist(turn.ZScore.H)
#MetaHit-Den STD = 0.4038558
keepmetahit <- which(data1$dataset == "MetaHit")
metahitdata <- data1[keepmetahit, ]
metahit.ZScore.H <- metahitdata$H.corr/0.4038558
hist(metahit.ZScore.H)
#Korean STD = 0.5400615
keepKorean <- which(data1$dataset == "Korean")
Koreandata <- data1[keepKorean, ]
Korean.ZScore.H <- Koreandata$H.corr/0.5400615
hist(Korean.ZScore.H)
#HMP STD = 0.6774514
keepHMP <- which(data1$dataset == "HMP")
hmpdata <- data1[keepHMP, ]
HMP.ZScore.H <- hmpdata$H.corr/0.6774514
hist(HMP.ZScore.H)
#CCHC STD = 0.5784522
keepCCHC <- which(data1$dataset == "CCHC")
CCHCdata <- data1[keepCCHC, ]
CCHC.ZScore.H <- CCHCdata$H.corr/0.5784522
hist(CCHC.ZScore.H)
#GLNE07 STD = 0.4663871
keepGLNE <- which(data1$dataset == "GLNE")
GLNEdata <- data1[keepGLNE, ]
GLNE.ZScore.H <- GLNEdata$H.corr/0.4663871
hist(GLNE.ZScore.H)
#Escobar STD = 0.5378935
keepescobar <- which(data1$dataset == "Escobar")
escobardata <- data1[keepescobar, ]
Escobar.ZScore.H <- escobardata$H.corr/0.5378935
hist(Escobar.ZScore.H)
#COMBO STD = 0.6870763
keepCOMBO <- which(data1$dataset == "COMBO")
COMBOdata <- data1[keepCOMBO, ]
COMBO.ZScore.H <- COMBOdata$H.corr/0.6870763
hist(COMBO.ZScore.H)
#COMBO STD = 0.3489582
keeptwinsUK <- which(data1$dataset == "twinsUK")
twinsUKdata <- data1[keeptwinsUK, ]
twinsUK.ZScore.H <- twinsUKdata$H.corr/0.3489582
hist(twinsUK.ZScore.H)

#data1 order Escobar, GLNE, CCHC, HMP, Korean, MetaHit, COMBO, twinsUK
HZScore <- c(Escobar.ZScore.H, GLNE.ZScore.H, CCHC.ZScore.H, HMP.ZScore.H, Korean.ZScore.H, metahit.ZScore.H, COMBO.ZScore.H, twinsUK.ZScore.H)
data1 <- cbind(HZScore, data1)

amish.data <- cbind(amish.ZScore.H, amish.data)
turn.data <- cbind(turn.ZScore.H, turn.data)
rm(Escobar.ZScore.H, GLNE.ZScore.H, CCHC.ZScore.H, HMP.ZScore.H, Korean.ZScore.H, metahit.ZScore.H, amish.ZScore.H, turn.ZScore.H, COMBO.ZScore.H, twinsUK.ZScore.H)
rm(keepmetahit, keepCCHC, keepescobar, keepGLNE, keepHMP, keepKorean)

@@

group.data.test1 <- cbind(comb1.H.Zscore, comb1.obese, comb1.bmi.class)
group.data.test2 <- cbind(comb2.H.Zscore, comb2.obese)
colnames(group.data.test1) <- c("H.Zscore", "obese", "BMI.class")
colnames(group.data.test2) <- c("H.Zscore", "obese")
#Look at Obese versus non-obese
t.test(H.Zscore ~ obese, data=group.data.test2) #P-value = 0.00493
table(group.data.test2$obese) # No = 1099 | Yes = 443
p1 <- ggplot(group.data.test2, aes(obese, H.Zscore)) + geom_boxplot()

#A more in-depth look, breaking down the groups
anova(lm(H.Zscore ~ BMI.class, data=group.data.test1)) #P-value = 0.03826
a2 <- aov(H.Zscore ~ BMI.class, data=group.data.test1)
posthoc <- TukeyHSD(x=a2, 'BMI.class', conf.level=0.95)
  # Obese-Lean (p-value=0.107), Over-Norm (p-value=0.872), Over-Obese (p-value=0.0394479)

table(group.data.test1$BMI.class) 
  # Lean = 509 | Overweight = 365 | Obese = 461
ddply(group.data.test1, ~BMI.class, summarise, mean=mean(H.Zscore), sd=sd(H.Zscore))
  # Lean = 0.027 +\- 0.993| Overweight = 0.059 +/- 0.982 | Obese = -0.111 +\- 1.016

@@

summary(lm(HZScore ~ bmi, data=data1)) #P-value 0.1272, R2= 0.001959
plot(HZScore ~ bmi, data=data1)
abline(lm(HZScore ~ bmi, data=data1), col="red")

@@

#Amish SD = 0.7918527
am.BF.corr <- as.character(amish.data$BF.ratio.corr)
am.BF.corr <- as.numeric(am.BF.corr) # Value 19 has no value
amish.norm.BF <- am.BF.corr/0.7918527
hist(amish.norm.BF)
#Turnbaugh SD = 0.4578937
turn.norm.BF <- turn.data$BF.corr/0.4578937
hist(turn.norm.BF)
#MetaHit SD = 0.9707577
keepmetahit <- which(data2$dataset == "MetaHit")
metahitdata <- data2[keepmetahit, ]
metahit.norm.BF <- metahitdata$BF.ratio.corr/0.9707577
hist(metahit.norm.BF)
#Korean SD = 0.6048195
keepKorean <- which(data2$dataset == "Korean")
Koreandata <- data2[keepKorean, ]
Korean.norm.BF <- Koreandata$BF.ratio.corr/0.6048195
hist(Korean.norm.BF)
#HMP SD = 8.301864
keepHMP <- which(data2$dataset == "HMP")
hmpdata <- data2[keepHMP, ]
HMP.norm.BF <- hmpdata$BF.ratio.corr/8.301864
hist(HMP.norm.BF)
#CCHC SD = 2.00527
keepCCHC <- which(data2$dataset == "CCHC")
CCHCdata <- data2[keepCCHC, ]
CCHC.norm.BF <- CCHCdata$BF.ratio.corr/2.00527
hist(CCHC.norm.BF)
#GLNE07 SD = 0.4451074
keepGLNE <- which(data2$dataset == "GLNE")
GLNEdata <- data2[keepGLNE, ]
GLNE.norm.BF <- GLNEdata$BF.ratio.corr/0.4451074
hist(GLNE.norm.BF)
#Escobar SD = 0.4232106
keepescobar <- which(data2$dataset == "Escobar")
escobardata <- data2[keepescobar, ]
Escobar.norm.BF <- escobardata$BF.ratio.corr/0.4232106
hist(Escobar.norm.BF)
#COMBO SD = 5.928664
keepCOMBO <- which(data2$dataset == "COMBO")
COMBOdata <- data2[keepCOMBO, ]
COMBO.norm.BF <- COMBOdata$BF.ratio.corr/5.928664
hist(COMBO.norm.BF)
#twinsUK SD = 0.9802656
keeptwinsUK <- which(data2$dataset == "twinsUK")
twinsUKdata <- data2[keeptwinsUK, ]
twinsUK.norm.BF <- twinsUKdata$BF.ratio.corr/0.9802656
hist(twinsUK.norm.BF)

#data1 order Escobar, GLNE, CCHC, HMP, Korean, MetaHit
BFScore <- c(Escobar.norm.BF, GLNE.norm.BF, CCHC.norm.BF, HMP.norm.BF, Korean.norm.BF, metahit.norm.BF, COMBO.norm.BF, twinsUK.norm.BF)
data2 <- cbind(BFScore, data2)

amish.data <- cbind(amish.norm.BF, amish.data)
turn.data <- cbind(turn.norm.BF, turn.data)
rm(Escobar.norm.BF, GLNE.norm.BF, CCHC.norm.BF, HMP.norm.BF, Korean.norm.BF, metahit.norm.BF, amish.norm.BF, turn.norm.BF, COMBO.norm.BF, twinsUK.norm.BF)
rm(keepmetahit, keepCCHC, keepescobar, keepGLNE, keepHMP, keepKorean)

@@

#Look at Obese versus non-obese
t.test(BF.ratio.corr ~ obese, data=group.data.test2) 
  # P-value = 0.07283
table(group.data.test2$obese)
  # No = 1099 | Yes = 443
bf1 <- ggplot(group.data.test2, aes(obese, BF.ratio.corr)) + geom_boxplot()

#A more in-depth look, breaking down the groups
anova(lm(BF.ratio.corr ~ BMI.class, data=group.data.test)) 
  # P-value = 0.005442
a2 <- aov(BF.ratio.corr ~ BMI.class, data=group.data.test)
posthoc <- TukeyHSD(x=a2, 'BMI.class', conf.level=0.95)
  # Obese-Norm (p-value=0.016) | Over-Norm (p-value=0.017), Over-Obese (p-value=0.979)
table(group.data.test$BMI.class)
  # Lean = 326 | Overweight = 240 | Obese = 262
ddply(group.data.test, ~BMI.class, summarise, mean=mean(BF.ratio.corr), sd=sd(BF.ratio.corr))
  # Lean = -0.392 +\- 1.319 | Overweight = -0.695 +/- 1.705 | Obese = -0.718 +\- 2.180
bf2 <- ggplot(group.data.test, aes(BMI.class, BF.ratio.corr)) + geom_boxplot()

summary(lm(BFScore ~ bmi, data=data2)) #P-value 0.06589, R2= 0.002847
plot(BFScore ~ bmi, data=data2)
abline(lm(BFScore ~ bmi, data=data2), col="red")

@@

###################################Bacteroidetes#######################################
#Look at Obese versus non-obese
t.test(Zbacter ~ obese, data=group.data.test2) 
  # P-value = 0.2892
table(group.data.test2$obese)
  # No = 1099 | Yes = 443
bacter1 <- ggplot(group.data.test2, aes(obese, Zbacter)) + geom_boxplot()
